---
title: "Incarceration Trends in the U.S. 1970 - 2018"
author: "Trevor Tang"
date: '2022-11-01'
output: html_document
---

```{r echo=FALSE, warning = FALSE, eval = TRUE, message = FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
source("../source/summary.R")
```

## Abstract 
[TODO: Include a short abstract (about 50 words)]

## Introduction 
The social issues related to incarcerating trends which this dataset and Vera
wants to tackle are those of mass incarceration. Thus, the data provided showcases
the rates and prison populations of people from the binary genders, and of a spectrum
of races. Analyzing this dataset might help decriminalize races that might
be disproportionally large in the prison system. Many sentences might also be
given for minor offenses. Vera wants to show that those of color, in poverty,
and our unjust prosecution system result in racial disparities in policing and
sentencing.

The direct stakeholders are the U.S. Justice system and their corresponding 
departments and employees as they will be able to visualize the trends of who
is being incarcerated (perhaps unjustly) into the prison system. Some central 
values that will be engaged by these people are fairness, human dignity, and
justice. An indirect stakeholder might be communities that are currently facing
overpolicing. Vera indicates that many prisoners are those police tend to target
more such as the homeless, people with mental illnesses, substance users, or
people of color. As such justice, privacy, and security will come to these 
people as well.

Some questions answered in this report are as follows: <br>
- 1. How has the U.S. jail population changed from 1978 - 2018 and why? <br>
- 2. How has the U.S. Jail population changed in the most populous states (CA, NV, FL, NY) and how does this compare with Washington? What conclusions can we draw from this? <br>
- 3. What does a map showing county's ratio of black prisoners to white prisoners reveal about geographic inequalities related to our prison systems?

## Data Summary 
IMPLEMENT

## Growth of U.S. Prison Population
```{r, echo = FALSE, fig.align = "center", fig.width = 10}
plot_jail_pop_for_us()
```
_Figure 1_: Increase of Jail Population in U.S. (1970 - 2018). This chart shows over time the population of jails has been increasing at a rather steady rate and then seeming to decline from 2008 onwards.

A key pattern in this graph is that while the jail population has been growing steadily, it had been declining since 2008. It might be that laws have been passed that created a more equitable justice system, or that prison reformation has resulted in prisoners being released more than they are being admitted. 

## Growth of Prison Population By State

```{r, echo = FALSE, fig.align = "center", fig.width = 10}
plot_jail_pop_by_states(c("CA", "NY", "TX", "FL", "GA", "WA"))
```
_Figure 2_: Increase of Jail Population of Certain States in the U.S. (1970 - 2018). States were chosen by most populous, and then Washington was included for comparison (and because this is for INFO 201 at the University of Washington). It is similar to Figure 1, but instead of having jail populations of the nation over time it is of a respective state instead.

A key pattern identified in this graph is how California has the highest jail population rate while Washington has the lowest. This might be because California is the most populated state or in general has more criminal activity. Another pattern is that all of these states have declining population rates around 2010. This is similar to _figure 1_ and might be based off the results of better prison reformation. 

## Ratio of Black Jail Populations to White Jail Populations
```{r, echo = FALSE, message = FALSE, fig.width = 20, fig.height = 10}
create_inequality_map()
```
_Figure 6_: Comparing the ratio of black jail populations to white jail populations per county in the Southern U.S. For example, a ratio of _6_ indicates that there are 6 black prisoners per every 1 white prisoner. This part of the United States was chosen since it contained ratios that were much greater than other parts of the United States.

**Note:** The population variables used in the above figure for the ratios were measured for a singular day and are not a summation for all the prisoners within a timespan such as a year. As such, we can still draw meaningful conclusions from it.

On closer observation, this map reveals a potential inequality such that black people seem to be put in prison more than white people. For instance, the greatest ratios found
in the Southern U.S. are in Coahoma, MS, Tunica, MS, and Clarendon, SC with the values of 104, 65, and 52 respectively. A more populous prison county, Shelby, TN, with a 3,883 black prison population and a 883 white prison population shows that for about every 4.5 black prisoners there is 1 white prisoner. As such, deep southern states, especially Mississippi and Alabama, seem to have an overwhelming number of black prisoners compared to white prisoners compared to the rest of the Southern U.S. 

It is rather concerning as in the counties listed above, the population breakdown is as follows:

```{r, echo = FALSE, message = FALSE}

knitr::kable(get_southern_populations)
```
While the black population outnumbers the white population, it still doesn't explain why the ratio of black prisoners to white prisoners is so high. It questions whether black neighborhoods are being overpoliced by officers in power, or whether these states are issues where blacks are disproportionately taking up prison space since they outnumber the white population.

